\chapter*{Acknowledgements}

First of all, I would like to thank my advisor, Marcelo de Oliveira Johann, who not only helped me a lot with the many specific details of sound synthesis and music theory but also was always available to help and patiently answered the many questions I had about this art I love but still don't understand as much as I wanted.

To Francisco Knebel, Rodrigo Neves and Leonardo Costa, classmates and dear friends without whom I definitively wouldn't be able to achieve the title of engineer. To Alexandre Saccol, Lucas Flores and Cassio Fachinelli who helped me in the many times computers got me confused and lost. To Dirceu Bueno, Lucas Jantsch, Pedro Fetter, Luci Bongiorni and Yi Chen Wu who now I'll finally be able to call professional colleagues. To Matheus Perius, who made the task of working full time and studying less of a pain.

To Maria, Milton, Guilherme and all my family, who gave me unconditional support to study and focus in what was important. To Martina, who were my safe haven and gave me strength throughout the hardest part of my journey to become the one thing I wanted to become since childhood: An inventor who is capable of creating things that sing and fly.

Finally, to Bob Miller and the many anonymous in the many different communities and forums spread across the entire globe that helped me answer many questions and guide me to wisdom. You all are the reason I still believe in a better society through cooperation and kindness.

\clearpage