\chapter{Introduction}
\label{cap:intro}

\section{The initial problem}

In order to navigate the real word, autonomous drones need to estimate their position
with reference to the world, with a certain precision \cite{introLivro1}. When used in
rural areas, drones are often used to disperse seeds, pesticides, or survey the field.
Field surveillance and inspection can now be done remotely, without the need for the
farmer to traverse the entire plantation, which could be thousands of acres of land
 \cite{introArtigo1}. Urban environments also benefit from drone services, from preventive
 maintenance services \cite{droneurban1} to emergency reponse infrastructure \cite{droneurban2}
\cite{droneurban3}.

\gls{gnss} is the main positioning system used for drones \cite{introLivro2}. Although GNSS
systems allow for more precise flight procedures, they come with a drawback: More precise and
robust systems usually involve heavier and bulkier hardware. Absolute positioning may not be
necessary in most cases though: Sometimes, the position of the drone is not as critical as its
position relative to a ground mission target, sometimes called \gls{poi}. While the accuracy of
the absolute position of the drone can be incresed over time through the use of filtering,
quick and precise relative positioning between the drone and ground target is critical so the drone
can either approach it or avoid it.

The purpose of this dissertation is to investigate the feasibility of a sensor fusion system for precise
ground target location, comprising of one distance measuring pod, to be attached to the drone as payload.
The drone system itself is out of the scope of this work.

\section{Mission scope}

As mentioned before, the drone absolute position acquisition system is not part of the scope of this
work. This dissertation will focus on the payload (also refered as pod) of the drone, to be connected
to the drone in order to measure distances. A first system will be developed to behave as a drone, and
through the use of computer vision will identify a test target and calculate it's direction relative
to the drone. This direction information will be transmitted to the pod, as a set of angles, and the
pod will then measure the distance between the pod and target. It will also compensate for parallax, 
as the pod could be positioned anywhere in the drone, and then estimate the distance between the
camera to the object, as if the camera itself had measured it instead of an external module.

Different sensors have different specifications. In order to investigate the many devices to design
a precise drone location system, well-defined requirements are of the utmost importance. As this
survey was carried out in Brazil, all system specifications will be defined by
Brazilian regulations and plantation standards.

The first parameter to define is the altitude range in which the drone will fly. Brazil's Ministry of
Agriculture defines, as of November 2024 \cite{dronealturaagro}, that the maximum flight altitude allowed for
for a drone used for agricultural purposes or far from crowded areas at is 400 feet (approximately 120 meters).
In addition to this resolution, the \gls{anac} defines that any drone that flies above 400
feet must be registered and licensed \cite{dronealturaANAC}. This is, of course, the maximum allowed
altitude of operations, and not necessarily the altitude at which the drone will, in fact, operate.

\section{System Architecture}

This theoretical target location system will consist on a target-to-drone distance measuring system.
The pod should first communicate to the drone system in order to receive orientation. Then measure the
distance to target on ground. To get correct distance (and thus position) of the target to the drone,
the pod system should make some corrections necessary for the difference in point-of-view.


\section{Different technologies for target location}

With the drone position measured with precision, the POI's position can be estimated by simple
geometry: The drone's position is a known point in space, and the system could measure the
distance between the drone and the target and the angle between the drone's orientation and the
observed POI.

\subsection{Ultrasonic Sensors}

Ultrasonic sensors consist of a sensor that emits ultrasonic pulses (above 20 kHz)
and measures the time between the emission and the receiving of an echo \cite{ultrasound1}.
Ultrasonic sensors are commonly used in robotics, automotive parking systems, and level detection
because of their ability to detect objects without requiring physical contact. They are inexpensive,
work well under various lighting conditions, and are effective in detecting both solid and liquid
surfaces, although their accuracy can be affected by environmental factors such as temperature,
air pressure, and the type of surface to be detected. Ultrasonic sensors rarely measure distances
greater than 10 meters \cite{ultrasound2}.

\subsection{Light-based sensors}

Similarly to radio waves, there are equivalent sensors that utilize electromagnetic radiation
in the visible spectrum. The main types of light-based sensors are the following
 \cite{infrared}\cite{laserrangefinder}:

\begin{itemize}
  \item \gls{tof} : These sensors emit light (typically laser or infrared) and measure
  the time it takes for the reflected light to return to the sensor. They are widely used in
  smartphones, robotics, and drones because of their precision and ability to operate over a range
  of distances. These modules can be multi-point or single-point of measurement.
  \item Light Detection and Ranging (LIDAR): These use laser beams to measure distance by timing the
  reflected light. LIDAR systems are commonly used in autonomous vehicles, mapping, and industrial
  applications due to their high accuracy and ability to create 3D maps. Usually consist of a light
  transmitter and receiver module rotating, in order to get wide angles of measurements.
  \item Structured Light Sensors: These sensors project a known light pattern (for example, a grid)
  onto an object and analyze the deformation of the pattern to calculate distance. They are used in
  depth-sensing cameras like Microsoft Kinect for 3D scanning and motion tracking.
  \item Stereo Vision: This method uses two or more cameras to simulate human binocular vision.
  By analyzing the disparity between images captured by each camera, it calculates the depth and distance.
  Stereo vision is used in robotics, AR/VR, and object recognition systems.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% tof + gimbal vs matrix tof vs lidar

TOF and LIDAR are preferred in the field of drones and other autonomous robots for their simplicity,
size, and wide range of specifications CITATION. Structured Light Sensors are more commonly used in
civil engineering, as they are usually too bulky to be fitted on a drone. Stereo visio is by far the
most complex of all the methods cited, being very complex and hard to calibrate CITATION.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Radar}

Radar modules are increasingly used in drones for applications such as obstacle detection,
collision avoidance, and terrain mapping \cite{radaraltimetro2}. By emitting radio waves and
analyzing the reflected signals, radar sensors provide precise distance and velocity measurements,
even in low visibility conditions such as fog, rain, or darkness, where optical sensors can
struggle \cite{introtoradar}. Radar's ability to detect objects over longer distances and through
cluttered environments makes it ideal for autonomous navigation and enhancing flight safety.
In addition, advanced radar systems, such as millimeter-wave radar, are used for altimetry and
precise positioning during landings, particularly in drones operating in challenging outdoor environments. 

Although radar modules offer many advantages for drones, they also have notable disadvantages.
One key limitation is their high power consumption, which can reduce the drone's battery life,
particularly in smaller systems. Additionally, radar systems can struggle with resolution limitations,
especially for small or closely spaced objects, making them less effective for detecting fine details
compared to optical sensors like LIDAR or cameras. Environmental factors, such as electromagnetic
interference, can also affect performance and introduce noise into the radar signal. Furthermore,
cost can be a barrier, as advanced radar systems, such as millimeter-wave radars, tend to be more
expensive than simpler alternatives like ultrasonic or infrared sensors. Lastly, the complexity of
data processing and integration with other onboard systems requires advanced algorithms and additional
computational resources, which may complicate the drone's design.

\section{Target location}

%justificativa da escolha de sensores e ranges de operacao

Of all types of sensors analyzed in this work, optical ranging sensors are the more capable and
precise for weight, volume, and price. Light-based sensors, such as infrared or laser rangefinders,
fit perfectly within the technical constraints of the design. Radar is too complex, bulky, and
expensive. Ultrasound sensors are simply not precise enough.

As the angle of the target relative to the drone module would be known via the camera, the only
measure required would be the distance of the target to the module. The rangefinder would be
measuring this distance from above, and the geometry errors corrected.

